Cooperative Jahn-Teller effect with rigid octahedral rotations in perovskites

### 2023-09-14 manuscript/paper.tex
full lattice vs AB cages, comparison fig expalined.

In a cubic perovskite ABO$_3$,
each transition metal atom/ion B
is surrounded by six O$^{2-}$, each at a distance $a/2$,
 making an octahedron of size $a$ around it.
The potential of these ions $V_{O}(\vec r)= V_{oct}(\vec r)$
splits the degenerate d-orbitals
by an amount ${\Delta_O=160q_oD_{4}/3a}$
 into $t_{2g}$ and $e_g$.
However, this is not the only crystal field potential at the B site.
The coordination cages of next nearest neighbours, 
Eight A atoms/cations all at the same distance $\sqrt{3}a/2$
make a cube around the B ion,
which also produces an octahedral field, given by 
\begin{align}
V_{A}(\vec r) = \frac{8}{81 \sqrt{3}}\frac{q_A}{q_o} V_{O}(\vec r).
\end{align}
Furthermore,
the nearest neighbouring B ions, size of them each at a distance $a$,
make an octahedron, with a potential 
\begin{align}
V_{B}(\vec r) = -\frac{1}{32}\frac{q_B}{q_o}V_{O}(\vec r).
\end{align}
It is interesting to note that 
$V_{A}(\vec r)$ and $V_{B}(\vec r)$ have opposite signs
and 
their combined effect 
$V_{AB}(\vec r) = V_{A}(\vec r)+ V_{B}(\vec r)$
depends on their relative charges.
Using the charge neutrality $q_B=3q_o-q_A$,
we can write 
\begin{align}
V_{AB}(\vec r) &= f_{q_A}V_{O}(\vec r),\\
f_{q_A} &= \left(\frac{1}{32}+\frac{8}{81 \sqrt{3}}\right)\frac{q_A}{q_o} -\frac{3}{32},
\end{align}
which shows that $V_{AB}(\vec r)\to 0$
as
${q_A/q_0\to 1458/({256 \sqrt{3}+243})\simeq 2.124}$.
%where they cancel each other exactly.

@@

whereas $V_{lm}^{B}$ is given by,

@@

Ignore the coupling between the $e_{g}$ and $t_{2g}$ manifolds (between $xy$ and $x^2-y^2$ at $\alpha=0$,
their Hamiltonians turn out to be diagonal.
We can write the single particle energies of $t_{2g}$ states
as
\begin{align}
\label{eq:theta0-Es}
(E_{xy},&E_{yz},E_{zx})= (2, -1, -1)
\left[\frac{}{} \Delta_{2}  D_2 + \Delta_4 D_4\right],
 \end{align}
 where $E_{yz}$ and $E_{zx}$ are degenerate and 
\begin{align}
\Delta_{l} =&  \left(q_A \Delta_{l}^A + q_B\Delta_{l}^B\right)/a,~(l=2,4),\\
\Delta_{2}^A =& 
%from A
\frac{64 \sin ^2\beta }{7 (\cos 2 \beta +2)^{5/2}} % simplifies.... from:
 %\frac{16 (3 \sin \beta+\sin 3\beta)^2}{7 (\cos 2\beta+2)^{9/2}},\\
 \simeq \frac{64}{63 \sqrt{3}}\beta ^2,\\
\Delta_4^A =&
 \frac{160 \sin ^2\beta}{63 (\cos 2\beta+2)^{9/2}} \times \\
&  (125 \cos 2\beta+42 \cos 4\beta+7 \cos 6\beta+82),\\
\simeq& \frac{40960}{5103 \sqrt{3}}\beta ^2,\\
% \Delta_4^A =&
% \frac{160 \sin ^2\beta (125 \cos 2\beta+42 \cos 4\beta+7 \cos 6\beta+82)}{63 (\cos 2\beta+2)^{9/2}},\\
 %from B
 \label{eq:theta0-delB2}
\Delta_{2}^B =& \frac{2}{7} \left(1-\sec ^3\beta\right)
\simeq -\frac{3}{7}\beta ^2,\\
\label{eq:theta0-delB4}
  \Delta_4^B =&\frac{5}{126} \left((7 \cos 4\beta-3) \sec ^5\beta-4\right)
  \simeq -\frac{115}{63}\beta ^2.
 %\left( \Delta_{2}  D_2 + \Delta_4 D_4\right),\\
 \end{align}
 We see that $\Delta_{2/4}^A>0$ while $\Delta_{2/4}^B<0$,
 revealing a competition between the two terms 
 that would determine the sign and size of the term
  in the square bracket in Eq.~\ref{eq:theta0-Es} 
%  (hence the ground state and splitting) 
  as $q_A$ (and hence $q_B$)
  is varied.
The average $t_{2g}$ energy
 also shifts from
$-2/5\Delta$ by an amount
%\rev{E0t2g or change in it? how to define the expressions... plotted is always change in it...}
\begin{align}
\label{eq:theta0-E0}
E_{0}^{t_{2g}} =& \left(q_A E_{0}^A + q_BE_{0}^B\right)D_4/a
-(2/5)f_{q_A}\Delta_{O},\\
E_{0}^A =&
 \frac{8 (92 \cos 2\beta+24 \cos 4\beta+20 \cos 6\beta+5 \cos 8\beta+51)}{9 (\cos 2\beta+2)^{9/2}},\\
 \simeq& \frac{512}{243 \sqrt{3}}-\frac{2560 \beta ^2}{729 \sqrt{3}},\\
E_{0}^B=&
 -\frac{1}{18} \left((5 \cos 4\beta+3) \sec ^5\beta+4\right)
 \simeq -\frac{2}{3}+\frac{10 \beta ^2}{9}.
 \end{align}

The results in Eqs.~\ref{eq:theta0-Es},\ref{eq:theta0-E0}
are summarised in Fig.~\ref{fig:t2g-alpha0},
where $E_{xy}$, $E_{yz/zx}$ and $E_0^{t_{2g}}$ are shown
as a function of $\beta$
at three different values of $q_A/q_o=1/2,1,2$.
We have 
assumed $D_2=\sqrt{D_4}$ and
taken $\Delta_o=4q_o/15a$ throughout this paper,
which fixes the value of $D_4=1/200$, 
and corresponds to $\Delta= 2.073$ eV at 
$q_o=2$ e and $a=7.0$ bohr.
We present the energies as a percentage of $\Delta_o$
by calculating all quantities at the above values of $q_o,a$.
Fig.~\ref{fig:t2g-alpha0} shows that
the the competition between A and B cages
determines the ground state and the size of the splitting between the 
$t_{2g}$ states
$E_{xy}$, and $E_{yz/zx}$.
In Fig.~\ref{fig:t2g-alpha0}(a),
where $q_A/q_o=1/2$ ($q_B/q_o=5/2$),
the effect of B cage dominates.
Since it's negative (see Eqs.~\ref{eq:theta0-delB2},\ref{eq:theta0-delB4}),
it reverses the order in Eq.~\ref{eq:theta0-Es}
and $E_{ex}$ becomes the lower than $E_{yz/zx}$, by about $6\%$ of $\Delta_o$ at $15^\circ$.
When A cage dominates at large $q_A/q_o$,
as shown in Fig.~\ref{fig:t2g-alpha0}(c) for $q_A/q_o=2$,
$E_{yz/zx}$ are the single particle ground states (again with a splitting of $\sim6\%$ of $\Delta_o$ at $15^\circ$),
whereas in the net effect of both cages
is reduced in Fig.~\ref{fig:t2g-alpha0}(b)
that corresponds to $q_A/q_o=1$
and we only obtain a splitting of $\sim2\%$ of $\Delta_o$ at $15^\circ$
between $E_{xy}$ and $E_{yz/zx}$ levels.
Obviously, there will also be a point near 
$q_A/q_o=1$ where A and B completely 
cancel each others effect (not shown in Fig.~\ref{fig:t2g-alpha0}).
We also see that $E_0^{t_{2g}}$ shows slight negative shift to a slight positive shift
when as $q_A/q_o$ increases when going from Fig.~\ref{fig:t2g-alpha0}(a)
to Fig.~\ref{fig:t2g-alpha0}(c).
This is because the rotation suppresses the
effect of AB cages on the
gap $\Delta$ between $e_{g}$ and $t_{2g}$.
Since A supports O while B opposes it,
$\Delta$ slightly widens at small $q_A$ and shrinks at large $q_A$,
shifting $E_0^{t_{2g}}$ slightly down or up with $\beta$.

@@

In this case, the situation gets slightly complicated as
the couplings between the three states become finite. 
However, 
the Hamiltonian still has a simple structure, 
\begin{align} 
\tilde H = 
\left(
\begin{array}{ccc}
 E_{xy} & g & -g \\
 g & E_{yz} & \lambda  \\
 -g & \lambda  & E_{yz} \\
\end{array}
\right),
\end{align}
where the matrix elements have relatively long expressions and can be found in the appendix. An approximation at small $\alpha,\beta$ is given later in this section. 
$\tilde H$ can be simplified by rotating the bases in $\{\ket{yz},\ket{zx}\}$ space
to obtain bright and dark states
as follows.
\begin{align}
\ket{B} &= \left(\ket{yz} - \ket{zx}\right)/\sqrt{2},\\
\ket{D} &=  (\ket{yz} + \ket{zx})/\sqrt{2}.
\end{align}
The state $\ket{D}$ at energy ${E_D= E_{yz}+\lambda}$ decouples 
from the rest (so is already an eigenstate of $\tilde H$) 
and the state $\ket{B}$ at energy ${E_{yz}-\lambda}$ obtains the enhanced collective coupling $\sqrt{2} g$. 
\com{
\begin{align}
\tilde H_{Bright} = 
\left(
\begin{array}{cc}
E_{xy} & \sqrt{2} g \\
\sqrt{2} g & E_{yz}-\lambda
\end{array}
\right).
\end{align}
The Hamiltonian can be solved analytically.
}
The Hamiltonian can thus be solved analytically to
obtain the remaining two
eigenstates
\begin{gather}
%\ket{+} = \cos\gamma \ket{xy} + \sin\gamma \left(\ket{yz} - \ket{zx}\right)/\sqrt{2},\\
%\ket{-} = \sin\gamma \ket{xy} - \cos\gamma \left(\ket{yz} - \ket{zx}\right)/\sqrt{2},\\
\ket{+} = \cos\gamma \ket{xy} + \sin\gamma \ket{B},\\
\ket{-} = \sin\gamma \ket{xy} - \cos\gamma \ket{B},\\
\tan2\gamma = \sqrt{8}g/(E_{xy}-E_{yz}+\lambda)
\end{gather}
at 
energies
\begin{gather}
E_{\pm} = \frac{E_{xy}+E_{yz}-\lambda}{2}  \pm \sqrt{(E_{xy}-E_{yz}+\lambda)^2 + 8g^2 }.
%\delta = E_{xy}-e_z+\lambda.
\end{gather}

@@

$E_D, E_{\pm}$ relative to their average
and the weight of $xy$ state $W_{xy}$ in $E_{\pm}$,
as a function of $\alpha$ at $q_A/q_o=1/2,1,2$ (and $\Delta=15q_o/4a$)
are shown in Fig.~\ref{fig:t2g-beta0}.
The average shifts slightly with $\alpha,q_A$ by 
$E_0^{t_{2g}}= \left(E_{xy} + 2 E_{yz}\right)/3-(2/5)f_{q_A}\Delta_{O}$,
again reflecting 
a change in the gap $\Delta$,
and is also included in Fig.~\ref{fig:t2g-beta0}.
Apart from the mixing between the three states $xy,yz,zx$ in $\ket{\pm}$ at finite $\alpha$, 
Fig.~\ref{fig:t2g-beta0} shows a similar trend as in Fig.~\ref{fig:t2g-alpha0}
but with an interesting twist.
%Fig.~\ref{fig:evals-alpha} reveals another interesting point.
In contrast to $\alpha=0$ case discussed in the previous section,
$W_{xy}$ is the largest in the highest (lowest) energy state at small (large) $q_A/q_o$,
as shown in Fig.~\ref{fig:t2g-beta0}(a,c).
$\alpha$ and $\beta$ would thus tend to produce different ordering
(details in the next section).
The splittings are also abut the same magnitude, between $4$-$6\%$ at $\alpha=15^\circ$
at $q_A/q_o=1/2,2$ [Fig.~\ref{fig:t2g-beta0}(a,c)] and about $2\%$ at $q_A/q_o=1$ [Fig.~\ref{fig:t2g-beta0}(b)].
The mixing is sizeable but still far from perfect ($50/50\%$) even at the largest $\alpha$ shown ($W_{xy}=0.25$-$0.35$ at $20^\circ$).
% At unrealistically large $\alpha\sim30-40^\circ$, 
Overall, at finite $\alpha$,
the lowest energy state is either purely $xy/yz$,
or contains at least some fraction of these,
but it is never pure $xy$.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.5\linewidth]{/Users/maz/repos/perovskite/calc/analytical-structure-param/fig-data/wts-energy-vs-beta-alpha-10-width-row-beta-0-15}
\caption{$\alpha=10^\circ$ case: exact numerical results. solution of full 5x5}
\label{fig:t2g-ge-wts}
\end{center}
\end{figure}

@@

The $t_{2g}$ Hamiltonian in the general case $\alpha,\beta\neq 0$ cannot be solve analytically
 so there is no longer any advantage of ignoring the coupling between 
 the $e_{g}$ and $t_{2g}$ manifolds.
Even though their effect should be negligible due to a large gap $\Delta$ between the two manifolds,
here we keep these couplings 
and numerically solve
the full Hamiltonian.
Figure~\ref{fig:t2g-ge-Es}
shows the energies at $\alpha=10^\circ$
as a function of $\beta$
at $q_A/q_o=1,2$ and $\Delta_o=4q_o/15a$ again.
All states in the $t_{2g}$ mix up to make the eigenstates $\ket{E_i}, i=1,2,3$
at energies $E_i, i=1,2,3$.
\az{E0, can we show E0 only in the first fig and not all of them?}
At $\beta=0$, these results correspond to $\alpha=10^\circ$ 
in Figs.~\ref{fig:t2g-beta0}(a,c),
so the lowest state at $\beta=0$
here is $\ket{D}$ in Figs.~\ref{fig:t2g-ge-Es}(a)
while it is the highest in Figs.~\ref{fig:t2g-ge-Es}(b).
We see two anticrossings or level repulsion in either of Figs.~\ref{fig:t2g-ge-Es}(a,b)
in different orders.
In Figs.~\ref{fig:t2g-ge-Es}(a), 
the anticorssing between the highest two states occurs at smaller $\beta$,
while the situation is the opposite in Figs.~\ref{fig:t2g-ge-Es}(b) where 
it occurs between the lowest pair at smaller $\beta$.
From the results in the previous two sections,
we know that not only A and B but also $\alpha$ and $\beta$ tend to establish different orders.
To get more insight, the weights of these three $t_{2g}$ states in the three eigenstates 
% in Fig.~\ref{fig:t2g-ge-Es}
are shown in Fig.~\ref{fig:t2g-ge-wts}.
Let's first focus on the left column ($q_A/q_o=1/2$). 
Fig.~\ref{fig:t2g-ge-wts}(a)
shows that, while $\ket{E_1}\to \ket{D}$ as $\beta\to 0$,
$\ket{E_1}$ practically stays in $yz,xz$ subspace
for $\beta<7^\circ$ and it quickly becomes $xy$
as $\beta\alpha=10^\circ$.
Fig.~\ref{fig:t2g-ge-wts}(b,c)
show a complimentary behaviour 
between $xy$ and $yz,xz$,
where $\ket{E_3}$
transforms from almost pure $xy$ at $\beta=0$
to almost complete $yz,xz$ at $\beta\gtrsim\alpha=10^\circ$.
The right column
in Fig.~\ref{fig:t2g-ge-wts} ($q_A/q_o=2$),
shows the opposite behaviour 
about the ``resonance'' $\beta\sim\alpha$.
Fig.~\ref{fig:t2g-ge-wts}(b,e)
show an interesting aspect,
the $W_{xy}$ peaking in the middle state $\ket{E_2}$
around before it transfers between $\ket{E_1}$ and $\ket{E_2}$
around $\beta\sim\alpha$,
i.e., at $\beta$
between the two anticrossings in 
Fig.~\ref{fig:t2g-ge-Es},
which makes sense if we imagine the continuation of various energy surfaces/curves
in 
the spectra in Fig.~\ref{fig:t2g-ge-Es}
ignoring their anticorssings:
$xy$ starts from the top and goes down in 
Fig.~\ref{fig:t2g-ge-Es}(a),
while it starts from the bottom and goes up
in Fig.~\ref{fig:t2g-ge-Es}(b).

@@

2x2 system so analytical sol always possible...
either theta or phi 0 givens no couplings but just shifts and eg orbitals stays as the eigenstates....

@@

presentation and discussion of eg splitting...
combined plots for the three cases w.r.t  angles... as three  rows
and two q values on the opposite side of the crossover...
so 2x3 panel figure.

\begin{align}
(H_{x^2-y^2},&H_{z^2}) = (1,-1)
\left[\frac{}{} \Omega_{2}  D_2 + \Omega_4 D_4\right],\\
\Omega_{l} =&  \left(q_A \Omega_{l}^A + q_B\Omega_{l}^B\right)/a,~(l=2,4),\\
\Omega_{2}^A =&  -\frac{64 \left(\alpha ^2 \left(7 \beta ^2+3\right)-6 \beta ^2\right)}{189 \sqrt{3}},\\
\Omega_{4}^A =& \frac{640 \left(\alpha ^2 \left(\beta ^2+16\right)-32 \beta ^2\right)}{1701 \sqrt{3}},\\
\Omega_{2}^B =&  \frac{3}{14} \left(\alpha ^2 \left(3 \beta ^2+2\right)-4 \beta ^2\right),\\
\Omega_{4}^B =& \frac{1}{168} (-5) \left(\alpha ^2 \left(37 \beta ^2+46\right)-92 \beta ^2\right)
\end{align}

>>>>

In a cubic perovskite ABO$_3$,
each transition metal atom/ion B
is surrounded by six O$^{2-}$, each at a distance $a/2$,
 making an octahedron of size $a$ around it.
The potential of these ions $V_{O}(\vec r)= V_{oct}(\vec r)$
splits the degenerate d-orbitals
by an amount ${\Delta_O=160q_oD_{4}/3a}$
 into $t_{2g}$ and $e_g$.
However, this is not the only crystal field potential at the B site.
The coordination cages of next nearest neighbours, 
Eight A atoms/cations all at the same distance $\sqrt{3}a/2$
make a cube around the B ion,
which also produces an octahedral field, given by 
\begin{align}
V_{A}(\vec r) = \frac{8}{81 \sqrt{3}}\frac{q_A}{q_o} V_{O}(\vec r).
\end{align}
Furthermore,
the nearest neighbouring B ions, size of them each at a distance $a$,
make an octahedron, with a potential 
\begin{align}
V_{B}(\vec r) = -\frac{1}{32}\frac{q_B}{q_o}V_{O}(\vec r).
\end{align}
It is interesting to note that 
$V_{A}(\vec r)$ and $V_{B}(\vec r)$ have opposite signs
and 
their combined effect 
$V_{AB}(\vec r) = V_{A}(\vec r)+ V_{B}(\vec r)$
depends on their relative charges.
Using the charge neutrality $q_B=3q_o-q_A$,
we can write 
\begin{align}
V_{AB}(\vec r) &= f_{q_A}V_{O}(\vec r),\\
f_{q_A} &= \left(\frac{1}{32}+\frac{8}{81 \sqrt{3}}\right)\frac{q_A}{q_o} -\frac{3}{32},
\end{align}
which shows that $V_{AB}(\vec r)\to 0$
as
${q_A/q_0\to 729/({256 \sqrt{3}+243})\simeq 1.062}$.
%where they cancel each other exactly.

@@

whereas $V_{lm}^{B}$ are given by,

@@

Ignoring the coupling between the $e_{g}$ and $t_{2g}$ manifolds (between $xy$ and $x^2-y^2$ at $\alpha=0$),
their Hamiltonians turn out to be diagonal.
We can write the single particle energies of $t_{2g}$ states
as
\begin{align}
\label{eq:theta0-Es}
(E_{xy},&E_{yz},E_{zx})= (2, -1, -1)
\left[\frac{}{} \Delta_{2}  D_2 + \Delta_4 D_4\right],
 \end{align}
 where $E_{yz}$ and $E_{zx}$ are degenerate and 
\begin{align}
\Delta_{l} =&  \left(q_A \Delta_{l}^A + q_B\Delta_{l}^B\right)/a,~(l=2,4),\\
\Delta_{2}^A =& 
%from A
\frac{64 \sin ^2\beta }{7 (\cos 2 \beta +2)^{5/2}} % simplifies.... from:
 %\frac{16 (3 \sin \beta+\sin 3\beta)^2}{7 (\cos 2\beta+2)^{9/2}},\\
 \simeq \frac{64}{63 \sqrt{3}}\beta ^2,\\
\Delta_4^A =&
 \frac{160 \sin ^2\beta}{63 (\cos 2\beta+2)^{9/2}} \times \\
&  (125 \cos 2\beta+42 \cos 4\beta+7 \cos 6\beta+82),\\
\simeq& \frac{40960}{5103 \sqrt{3}}\beta ^2,\\
% \Delta_4^A =&
% \frac{160 \sin ^2\beta (125 \cos 2\beta+42 \cos 4\beta+7 \cos 6\beta+82)}{63 (\cos 2\beta+2)^{9/2}},\\
 %from B
 \label{eq:theta0-delB2}
\Delta_{2}^B =& \frac{2}{7} \left(1-\sec ^3\beta\right)
\simeq -\frac{3}{7}\beta ^2,\\
\label{eq:theta0-delB4}
  \Delta_4^B =&\frac{5}{126} \left((7 \cos 4\beta-3) \sec ^5\beta-4\right)
  \simeq -\frac{115}{63}\beta ^2.
 %\left( \Delta_{2}  D_2 + \Delta_4 D_4\right),\\
 \end{align}
 We see that $\Delta_{2/4}^A>0$ while $\Delta_{2/4}^B<0$,
 revealing a competition between the two terms 
 that would determine the sign and size of the term
  in the square bracket in Eq.~\ref{eq:theta0-Es} 
%  (hence the ground state and splitting) 
  as $q_A$ (and hence $q_B$)
  is varied.
The average $t_{2g}$ energy
 also shifts from
$-2/5\Delta$ by an amount
%\rev{E0t2g or change in it? how to define the expressions... plotted is always change in it...}
\begin{align}
\label{eq:theta0-E0}
E_{0}^{t_{2g}} =& \left(q_A E_{0}^A + q_BE_{0}^B\right)D_4/a
-(2/5)f_{q_A}\Delta_{O},\\
E_{0}^A =&
 \frac{8 (92 \cos 2\beta+24 \cos 4\beta+20 \cos 6\beta+5 \cos 8\beta+51)}{9 (\cos 2\beta+2)^{9/2}},\\
 \simeq& \frac{512}{243 \sqrt{3}}-\frac{2560 \beta ^2}{729 \sqrt{3}},\\
E_{0}^B=&
 -\frac{1}{18} \left((5 \cos 4\beta+3) \sec ^5\beta+4\right)
 \simeq -\frac{2}{3}+\frac{10 \beta ^2}{9}.
 \end{align}

The results in Eqs.~\ref{eq:theta0-Es},\ref{eq:theta0-E0}
are summarised in Fig.~\ref{fig:t2g-alpha0},
where $E_{xy}$, $E_{yz/zx}$ and $E_0^{t_{2g}}$ are shown
as a function of $\beta$
at three different values of $q_A/q_o=1/2,1,2$.
We have 
assumed $D_2=\sqrt{D_4}$ and
taken $\Delta_o=4q_o/15a$ throughout this paper,
which fixes the value of $D_4=1/200$, 
and corresponds to $\Delta= 2.073$ eV at 
$q_o=2$ e and $a=7.0$ bohr.
We present the energies as a percentage of $\Delta_o$
by calculating all quantities at the above values of $q_o,a$.
Fig.~\ref{fig:t2g-alpha0} shows that
the the competition between A and B cages
determines the ground state and the size of the splitting between the 
$t_{2g}$ states
$E_{xy}$, and $E_{yz/zx}$.
In Fig.~\ref{fig:t2g-alpha0}(a),
where $q_A/q_o=1/2$ ($q_B/q_o=5/2$),
the effect of B cage dominates.
Since it's negative (see Eqs.~\ref{eq:theta0-delB2},\ref{eq:theta0-delB4}),
it reverses the order in Eq.~\ref{eq:theta0-Es}
and $E_{xy}$ becomes lower than $E_{yz/zx}$, by about $6\%$ of $\Delta_o$ at $15^\circ$.
When A cage dominates at large $q_A/q_o$,
as shown in Fig.~\ref{fig:t2g-alpha0}(c) for $q_A/q_o=2$,
$E_{yz/zx}$ are the single particle ground states (again with a splitting of $\sim6\%$ of $\Delta_o$ at $15^\circ$),
whereas the net effect of both cages
is reduced in Fig.~\ref{fig:t2g-alpha0}(b)
that corresponds to $q_A/q_o=1$
and we only obtain a splitting of $\sim2\%$ of $\Delta_o$ at $15^\circ$
between $E_{xy}$ and $E_{yz/zx}$ levels.
Obviously, there will also be a point near 
$q_A/q_o=1$ where A and B completely 
cancel each others effect (not shown in Fig.~\ref{fig:t2g-alpha0}).
We also see that $E_0^{t_{2g}}$ changes from a slight negative shift to a slight positive shift
with an increase in $q_A/q_o$ as we
go from Fig.~\ref{fig:t2g-alpha0}(a)
to Fig.~\ref{fig:t2g-alpha0}(c).
This is because the rotation suppresses the
effect of AB cages on the
gap $\Delta$ between $e_{g}$ and $t_{2g}$.
Since A supports O while B opposes it,
$\Delta$ slightly widens at small $q_A$ and shrinks at large $q_A$,
shifting $E_0^{t_{2g}}$ slightly down or up with $\beta$.

@@

In this case, the situation gets slightly complicated as
the couplings between the three states become finite. 
However, 
the Hamiltonian still has a simple structure, 
\begin{align} 
\tilde H = 
\left(
\begin{array}{ccc}
 E_{xy} & g & -g \\
 g & E_{yz} & \lambda  \\
 -g & \lambda  & E_{yz} \\
\end{array}
\right),
\end{align}
where the matrix elements have relatively long expressions and 
only an approximation at small $\alpha,\beta$ is given later in this section.
%\com{are not given here \az{[?? can be found in the appendix.]} An approximation at small $\alpha,\beta$ is given later in this section.}
$\tilde H$ can be simplified by rotating the bases in $\{\ket{yz},\ket{zx}\}$ space
to obtain bright and dark states
as follows.
\begin{align}
\ket{B} &= \left(\ket{yz} - \ket{zx}\right)/\sqrt{2},\\
\ket{D} &=  (\ket{yz} + \ket{zx})/\sqrt{2}.
\end{align}
The state $\ket{D}$ at energy ${E_D= E_{yz}+\lambda}$ decouples 
from the rest (so is already an eigenstate of $\tilde H$) 
and the state $\ket{B}$ at energy ${E_{yz}-\lambda}$ obtains the enhanced collective coupling $\sqrt{2} g$. 
\com{
\begin{align}
\tilde H_{Bright} = 
\left(
\begin{array}{cc}
E_{xy} & \sqrt{2} g \\
\sqrt{2} g & E_{yz}-\lambda
\end{array}
\right).
\end{align}
The Hamiltonian can be solved analytically.
}
The Hamiltonian can thus be solved analytically to
obtain the remaining two
eigenstates
\begin{gather}
%\ket{+} = \cos\gamma \ket{xy} + \sin\gamma \left(\ket{yz} - \ket{zx}\right)/\sqrt{2},\\
%\ket{-} = \sin\gamma \ket{xy} - \cos\gamma \left(\ket{yz} - \ket{zx}\right)/\sqrt{2},\\
\ket{+} = \cos\gamma \ket{xy} + \sin\gamma \ket{B},\\
\ket{-} = \sin\gamma \ket{xy} - \cos\gamma \ket{B},\\
\tan2\gamma = \sqrt{8}g/(E_{xy}-E_{yz}+\lambda)
\end{gather}
at 
energies
\begin{gather}
E_{\pm} = \frac{E_{xy}+E_{yz}-\lambda}{2}  \pm \sqrt{(E_{xy}-E_{yz}+\lambda)^2 + 8g^2 }.
%\delta = E_{xy}-e_z+\lambda.
\end{gather}

@@

$E_D, E_{\pm}$ relative to their average
and the weight of $xy$ state $W_{xy}$ in $E_{\pm}$,
as a function of $\alpha$ at $q_A/q_o=1/2,1,2$ (and $\Delta=15q_o/4a$)
are shown in Fig.~\ref{fig:t2g-beta0}.
The average shifts slightly with $\alpha,q_A$ by 
$E_0^{t_{2g}}= \left(E_{xy} + 2 E_{yz}\right)/3-(2/5)f_{q_A}\Delta_{O}$,
again reflecting 
a change in the gap $\Delta$,
and is also included in Fig.~\ref{fig:t2g-beta0}.
Apart from the mixing between the three states $xy,yz,zx$ in $\ket{\pm}$ at finite $\alpha$, 
Fig.~\ref{fig:t2g-beta0} shows a similar trend as in Fig.~\ref{fig:t2g-alpha0}
but with an interesting twist.
%Fig.~\ref{fig:evals-alpha} reveals another interesting point.
In contrast to $\alpha=0$ case discussed in the previous section,
$W_{xy}$ is the largest in the highest (lowest) energy state at small (large) $q_A/q_o$,
as shown in Fig.~\ref{fig:t2g-beta0}(a,c).
$\alpha$ and $\beta$ would thus tend to produce different ordering
(details in the next section).
The splittings are also about the same magnitude, between $4$-$6\%$ at $\alpha=15^\circ$
at $q_A/q_o=1/2,2$ [Fig.~\ref{fig:t2g-beta0}(a,c)] and about $2\%$ at $q_A/q_o=1$ [Fig.~\ref{fig:t2g-beta0}(b)].
The mixing is sizeable but still far from perfect ($50/50\%$) even at the largest $\alpha$ shown ($W_{xy}=0.25$-$0.35$ at $20^\circ$).
% At unrealistically large $\alpha\sim30-40^\circ$, 
Overall, at finite $\alpha$,
the lowest energy state is either purely $xy/yz$,
or contains at least some fraction of these,
but it is never pure $xy$.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.5\linewidth]{/Users/maz/repos/perovskite/calc/analytical-structure-param/fig-data/wts-energy-vs-beta-alpha-10-width-row-beta-0-15}
\caption{
\az{Wts of eg in these states $\lesssim0.2\%$...} 
$\alpha=10^\circ$ case: exact numerical results. solution of full 5x5}
\label{fig:t2g-ge-wts}
\end{center}
\end{figure}

@@

The $t_{2g}$ Hamiltonian in the general case $\alpha,\beta\neq 0$ cannot be solve analytically
 so there is no longer any advantage of ignoring the coupling between 
 the $e_{g}$ and $t_{2g}$ manifolds.
Even though their effect should be negligible due to a large gap $\Delta$ between the two manifolds,
here we keep these couplings 
and numerically solve
the full Hamiltonian.
Figure~\ref{fig:t2g-ge-Es}
shows the energies at $\alpha=10^\circ$
as a function of $\beta$
at $q_A/q_o=1/2,2$ and $\Delta_o=4q_o/15a$ again.
All states in the $t_{2g}$ mix up to make the eigenstates $\ket{E_i}, i=1,2,3$
at energies $E_i, i=1,2,3$.
\az{E0, can we show E0 only in the first fig and not all of them?}
At $\beta=0$, these results correspond to $\alpha=10^\circ$ 
in Figs.~\ref{fig:t2g-beta0}(a,c),
so the lowest state at $\beta=0$
here is $\ket{D}$ in Figs.~\ref{fig:t2g-ge-Es}(a)
while it is the highest in Figs.~\ref{fig:t2g-ge-Es}(b).
We see two anticrossings or level repulsion in either of Figs.~\ref{fig:t2g-ge-Es}(a,b)
in different orders.
In Figs.~\ref{fig:t2g-ge-Es}(a), 
the anticorssing between the highest two states occurs at smaller $\beta$,
while the situation is the opposite in Figs.~\ref{fig:t2g-ge-Es}(b) where 
it occurs between the lowest pair at smaller $\beta$.
From the results in the previous two sections,
we know that not only A and B but also $\alpha$ and $\beta$ tend to establish different orders.
To get more insight, the weights of these three $t_{2g}$ states in the three eigenstates 
% in Fig.~\ref{fig:t2g-ge-Es}
are shown in Fig.~\ref{fig:t2g-ge-wts}.
Let's first focus on the left column ($q_A/q_o=1/2$). 
Fig.~\ref{fig:t2g-ge-wts}(a)
shows that, while $\ket{E_1}\to \ket{D}$ as $\beta\to 0$,
$\ket{E_1}$ practically stays in $yz,xz$ subspace
for $\beta<7^\circ$ and it quickly becomes $xy$
as $\beta\alpha=10^\circ$.
Fig.~\ref{fig:t2g-ge-wts}(b,c)
show a complimentary behaviour 
between $xy$ and $yz,xz$,
where $\ket{E_3}$
transforms from almost pure $xy$ at $\beta=0$
to almost complete $yz,xz$ at $\beta\gtrsim\alpha=10^\circ$.
The right column
in Fig.~\ref{fig:t2g-ge-wts} ($q_A/q_o=2$),
shows the opposite behaviour 
about the ``resonance'' $\beta\sim\alpha$.
Fig.~\ref{fig:t2g-ge-wts}(b,e)
show an interesting aspect,
the $W_{xy}$ peaking in the middle state $\ket{E_2}$
around before it transfers between $\ket{E_1}$ and $\ket{E_2}$
around $\beta\sim\alpha$,
i.e., at $\beta$
between the two anticrossings in 
Fig.~\ref{fig:t2g-ge-Es},
which makes sense if we imagine the continuation of various energy surfaces/curves
in 
the spectra in Fig.~\ref{fig:t2g-ge-Es}
ignoring their anticorssings:
$xy$ starts from the top and goes down in 
Fig.~\ref{fig:t2g-ge-Es}(a),
while it starts from the bottom and goes up
in Fig.~\ref{fig:t2g-ge-Es}(b).

@@

\com{
So we consider the general case of

@@

presentation and discussion of eg splitting...
combined plots for the three cases w.r.t  angles... as three  rows
and two q values on the opposite side of the crossover...
so 2x3 panel figure.
}
